---
title: "baker_demo -- demonstration of nested partially latent class model inference"
author: "Zhenke Wu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{baker_demo}
  %\VignetteEncoding{UTF-8}
output:
  rmarkdown::html_vignette:
    number_sections: yes
    toc: no
---

<!--
#######################################################################
# Simulated Examples for Nested Partially-Latent Class Models
# Zhenke Wu | zhenkewu@gmail.com
# Februrary 16, 2016
-->

```{r setup,echo=FALSE,results="hide"}
# suppressPackageStartupMessages({
library(rjags)
library(R2jags)
library(baker)
library(binom)
library(knitr)
# })
```

# Introduction

In this vignette, a sharply truncated sampling procedure
is used to demonstrate use of the baker package tools.

The data for the illustration are related to pathogen
categorization in pneumonia.
```{r lkd}
fname = file.path(
  "example_data", 
  "pathogen_category_simulation.csv")
fname = system.file(
  fname, 
  package = "baker")
demodat = read.csv(fname, stringsAsFactors = FALSE)
kable(head(demodat))
```



# Setup



```{r doit}
rm(list=ls())
# Note: the example will only run 100 Gibbs sampling steps to save computing time.
# To produce useful posterior inferences, please modify "mcmc_options" as follows
#                      "n.itermcmc" to 50000
#                      "n.burnin"   to 10000,
#                      "n.thin"     to 40,


working_dir <- tempdir() # <-- create a temporary directory.
#curd = getwd()
#randname = paste(curd, basename(tempfile()), sep="/") # need absolute path
#dir.create(randname)
#working_dir = randname

K.true  <- 2   # no. of latent subclasses in actual simulation. 
# If eta = c(1,0), K.true is effectively 1.
J       <- 6   # no. of pathogens.
N       <- 500 # no. of cases/controls.

# case subclass weight (five values):
subclass_mix_seq <- c(0,0.25,0.5,0.75,1)


NREP   <- 100
MYGRID <- expand.grid(list(rep   = 1:NREP, # data replication.
                           iter  = seq_along(subclass_mix_seq),# mixing weights.
                           k_fit = c(1,2), # model being fitted: 1 for pLCM; >1 for npLCM.
                           scn   = 3:1)    # index for different truth; see "scn_collection.R".
)

n_seed   <- nrow(unique(MYGRID[,-3]))
seed_seq <- rep(1:n_seed,times=length(unique(MYGRID[,3])))

SEG   <- 1 # The value could be 1 to nrow(MYGRID)=3000; here we just simulate one data set.
scn   <- MYGRID$scn[SEG]
k_fit <- 2#MYGRID$k_fit[SEG] 
iter  <- MYGRID$iter[SEG] 
rep   <- MYGRID$rep[SEG] 


# current parameters:
curr_mix <- subclass_mix_seq[iter]
lambda   <- c(0.5,0.5) #c(curr_mix,1-curr_mix)
eta      <- c(curr_mix,1-curr_mix) 

# set fixed simulation sequence:
seed_start <- 20161215  
set.seed(seed_start+seed_seq[SEG])

if (scn == 3){
  ThetaBS_withNA <- cbind(c(0.95,0.95,0.55,0.95,0.95,0.95),
                          c(0.95,0.55,0.95,0.55,0.55,0.55))
  PsiBS_withNA   <- cbind(c(0.4,0.4,0.05,0.2,0.2,0.2),
                          c(0.05,0.05,0.4,0.05,0.05,0.05))
}

if (scn == 2){
  ThetaBS_withNA <- cbind(c(0.95,0.9,0.85,0.9,0.9,0.9),
                          c(0.95,0.9,0.95,0.9,0.9,0.9))
  PsiBS_withNA   <- cbind(c(0.3,0.3,0.15,0.2,0.2,0.2),
                          c(0.15,0.15,0.3,0.05,0.05,0.05))
}

if (scn == 1){
  ThetaBS_withNA <- cbind(c(0.95,0.9,0.9,0.9,0.9,0.9),
                          c(0.95,0.9,0.9,0.9,0.9,0.9))
  PsiBS_withNA   <- cbind(c(0.25,0.25,0.2,0.15,0.15,0.15),
                          c(0.2,0.2,0.25,0.1,0.1,0.1))
}



# the following paramter names are set using names in the 'baker' package:
set_parameter <- list(
  cause_list      = c(LETTERS[1:J]),
  etiology        = c(0.5,0.2,0.15,0.05,0.05,0.05),# same length as cause_list.
  pathogen_BrS    = LETTERS[1:J],
  meas_nm         = list(MBS = c("MBS1")), # a single source of Bronze Standard (BrS) data.
  Lambda          = lambda,              #ctrl mix (subclass weights).
  Eta             = t(replicate(J,eta)), #case mix; # of rows equals length(cause_list).
  PsiBS           = PsiBS_withNA,
  ThetaBS         = ThetaBS_withNA,
  Nu      =     N, # control sample size.
  Nd      =     N  # case sample size.
)

# # visualize pairwise log odds ratios for cases and controls when eta changes 
# # from 0 to 1. In the following simulation, we just use one value: eta=0.
# example("compute_logOR_single_cause")

simu_out   <- simulate_nplcm(set_parameter)
data_nplcm <- simu_out$data_nplcm
```

# Exploratory data analysis

```{r expl, fig.height=10, fig.width=10}

# specify cause list:
cause_list <- set_parameter$cause_list

# specify measurements:
# bronze-standard measurements:
patho_BrS_MBS1      <- set_parameter$pathogen_BrS
BrS_object_1        <- make_meas_object(patho_BrS_MBS1,"MBS","1","BrS",cause_list)
   # please use ?make_meas_object to see the measurement standards.


# pairwise log odds ratio plot:
pathogen_display <- BrS_object_1$patho
plot_logORmat(data_nplcm,pathogen_display,1)
```


# Model specification

```{r domod}
m_opt1 <- list(likelihood   = list(cause_list = cause_list,               # <---- fitted causes.
                                   k_subclass = k_fit,                    # <---- no. of subclasses.
                                   Eti_formula = "~ 0",                   # <---- only apply FPR formula to specified slice of measurements; if not default to the first slice.
                                   FPR_formula = list(MBS1 = "~0")),      # <---- etiology regression formula.
               use_measurements = c("BrS"),                               # <---- which measurements to use to inform etiology
               prior        = list(Eti_prior   = overall_uniform(1, cause_list) ,                       # <--- etiology prior. 
                                   TPR_prior   = list(
                                     BrS  = list(info  = "informative",
                                                 input = "direct_beta_param",
                                                 val   = list(
                                                          MBS1 = list(alpha = list(rep(6,length(set_parameter$pathogen_BrS))),
                                                                      beta  = list(rep(2,length(set_parameter$pathogen_BrS)))
                                                                     )
                                                 )
                                                 
                                     )
                                     
                                   )# <---- TPR prior.
               )
)                       

model_options <- m_opt1
assign_model(model_options,data_nplcm)
```

# Fitting the model

```{r dofit}
# date stamp for analysis:
Date     <- gsub("-", "_", Sys.Date())
# include stratification information in file name:
dated_strat_name    <- file.path(working_dir, 
                                 paste0("scn_",scn,"_mixiter_",iter))
if (dir.exists(dated_strat_name)) {
  unlink(dated_strat_name, force = TRUE)
}

# create folder
dir.create(dated_strat_name)
fullname <- dated_strat_name

# for finer scenarios, e.g., different types of analysis applicable to the
# same data set. Here we just perform one analysis:
result_folder <- file.path(
  fullname,
  paste0("rep_", rep, "_kfit_",
         model_options$likelihood$k_subclass))
dir.create(result_folder)


# options for MCMC chains:
mcmc_options <- list(
  debugstatus = !TRUE,
  individual.pred = !TRUE,
  ppd             = TRUE,
  n.chains   = 1,
  n.itermcmc = 100, #50000
  n.burnin   = 50, #10000
  n.thin     = 1, #50
  result.folder = result_folder,
  bugsmodel.dir = result_folder,
  jags.dir = "",
  use_jags = TRUE
)

# Record the settings of current analysis:
# data clean options:
fname = file.path(
  "example_data", 
  "pathogen_category_simulation.csv")
fname = system.file(
  fname, 
  package = "baker")
global_patho_taxo_dir = fname

clean_options <- list(
  BrS_objects        =  make_list(BrS_object_1),         # <---- all bronze-standard measurements.
  patho_taxo_dir = global_patho_taxo_dir,
  allow_missing      = FALSE)

dput(clean_options,
     file.path(mcmc_options$result.folder,
               "data_clean_options.txt"))

gs <- nplcm(data_nplcm, model_options, mcmc_options)
```

# Visualizations

```{r dovizsetup}
result_folder <- mcmc_options$result.folder
```

## plot data, prior, and posteriors for all causitive pathogens:


```{r lkpan,fig.height=10, fig.width=10}
suppressWarnings(plot_panels(result_folder, bg_color=NULL))
```

## plot data, prior, and posterior for selected causes:

```{r lkpan2,fig.height=10, fig.width=10}
plot_panels(
  result_folder, bg_color = NULL, 
  select_latent = c("A","C"), exact = TRUE)
```

## plot the joint posterior of the etiology fraction for any three causes:

```{r trysel,fig.height=10, fig.width=10}
plot_selected_etiology(
  selected = c("A","B","C"), 
  result_folder, 1, 5)
```

## grouped etiology (e.g., bacteria (left) vs virus (right) )

```{r dodod,fig.height=10, fig.width=10}
plot_group_etiology(
  result_folder,
  dir_taxo = global_patho_taxo_dir, 
  ksFrac = 1)
```

## model checking by comparing observed pairwise log odds ratios (LOR)

 We compare observed LOR to the posterior predictive distributions of pairwise LOR; The numbers are
 (predicted LOR - observed LOR)/ s.e. (posterior predictive distribution of LOR);
 The closer to zero the better.

```{r doSLORD,fig.height=10, fig.width=10}
plot_check_pairwise_SLORD(result_folder, slice=1)
```

## model checking by comparing observed frequencies of binary patterns to the model-predicted ones

```{r doobsf,fig.height=10, fig.width=10}
dir_list <- as.list(c(result_folder))
plot_check_common_pattern(dir_list,slice_vec =c(1,1))
```

