---
title: "Manual for R Package: `nplcm`"
author: "Zhenke Wu <br> 
        [zhwu@jhu.edu](zhwu@jhu.edu) <br> 
        Department of Biostatistics <br>
        Johns Hopkins Bloomberg School of Public Health"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Variable and Function Definitions

We collect variable and function definitions below for easier reference. Currently, we use names that are motivated by PERCH study to provide an easy and scientifically
grounded interface for analysts. We will convert them to generic names in the framework of [nested partially latent
class models](http://biostats.bepress.com/jhubiostat/paper276/) in future versions of `nplcm`, because the model is a general discrete state model.

Raw Data
---
Inputs include (wrapped in `data_nplcm`):

* raw measurement data (lab tests, clinical information with case-control status, 
    other covaraite information)
* pathogen names
* pathogen taxonomy information 

Clean Data
---
* `clean_perch_data()`
    - `clean_combine_subsites()`
    - `extract_data_raw()`
    
Exploratory Data Analysis (EDA)
---
All EDA functions are in `eda.R`:

* `plot_logORmat()`

Choosing Model
---
* `assign_model()` chooses the model to fit based on user-specified model options
and the actual data supplied for analysis.

Fit Model
---

Check Model
---

Visualize Results
---

# Developer Notes

Pathogen orders:

1. raw pathogen orders input by the data
2. reorganized pathogen orders 
3. pathogen orders for display
















<!---
Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

-->
